### 2020-04-06 Computational Methods/01_intro-numerical-analysis.rmd
sln refactor + case studies/computational methods/statistical computing work
sln refactor + case studies/computational methods/statistical computing work

---
title: ''
mainfont: Arial
fontsize: 12pt
documentclass: report
header-includes:
- \PassOptionsToPackage{table}{xcolor}
- \usepackage{caption}
- \usepackage{amssymb}
- \usepackage{booktabs}
- \usepackage{longtable}
- \usepackage{array}
- \usepackage{multirow}
- \usepackage{wrapfig}
- \usepackage{float}
- \usepackage{colortbl}
- \usepackage{pdflscape}
- \usepackage{tabu}
- \usepackage{threeparttable}
- \usepackage{threeparttablex}
- \usepackage[normalem]{ulem}
- \usepackage{makecell}
- \usepackage[table]{xcolor}
- \usepackage{fancyhdr}
- \usepackage{boldline}
- \usepackage{tipa}
   \definecolor{headergrey}{HTML}{545454}
   \definecolor{msdblue}{HTML}{1C93D1}
   \pagestyle{fancy}
   \setlength\headheight{30pt}
   \rhead{\color{headergrey}\today}
   \fancyhead[L]{\color{headergrey}Moretz, Brandon}
   \fancyhead[C]{\Large\bfseries\color{headergrey}Geolocation}
   \rfoot{\color{headergrey}Chapter 1}
   \lfoot{\color{headergrey}}
   \fancyfoot[C]{\rmfamily\color{headergrey}Case Studies In Data Science}
geometry: left = 1cm, right = 1cm, top = 2cm, bottom = 3cm
date: "`r format(Sys.time(), '%d %B, %Y')`"
output:
  pdf_document:
    fig_caption: yes
    latex_engine: xelatex
editor_options: 
  chunk_output_type: console
---

@@

library(codetools, quietly = TRUE, warn.conflicts = FALSE)
library(lattice, quietly = TRUE, warn.conflicts = FALSE)
library(fields, quietly = TRUE, warn.conflicts = FALSE)

@@

# Geolocation

## Predicting Location via Indoor Positioning Systems

### The Raw Data

@@

file_offline <- file.path(data.dir, "offline.final.trace.txt")
file_online <- file.path(data.dir, "online.final.trace.txt")

raw_offline <- read_lines(file_offline)
raw_online <- read_lines(file_online)

@@

#### Sanity Check

```{r}
# number of comment lines in the data
sum(substr(raw_offline, 1, 1) == "#")

# total number of lines in the data
length(raw_offline)
```

#### Data Pre-processing

Generate read data function for preprocessing training and test data.

processLine <- function(x) {
   tokens <- strsplit(x, "[;=,]")[[1]]

   if(length(tokens) == 10)
      return(NULL)

   tmp <- matrix(tokens[ - (1:10)], ncol = 4, byrow = T)
   cbind(matrix(tokens[c(2, 4, 6:8, 10)], nrow = nrow(tmp),
                ncol = 6, byrow = T), tmp)
}

validLines <- function(data) {
   substr(data, 1, 1) != "#"
}

roundOrientation <- function(angles) {
   refs = seq(0, by = 45, length = 9)
   q <- sapply(angles, function(o) which.min(abs(o - refs)))
   c(refs[1:8], 0)[q]
}

readData <- function(file, submacs = macs) {

   lines <- read_lines(file)

   valid_lines <- lines[ validLines(lines) ]

   processed_lines <- lapply(valid_lines, processLine)

   data <- as.data.table(do.call("rbind", processed_lines),
                         stringsAsFactors = F)

   names(data) <- c("time", "scanMac", "posX", "posY", "posZ",
                       "orientation", "mac", "signal", 
                       "channel", "type")

   numVars <- c("time", "posX", "posY", "posZ",
                "orientation", "signal")

   data[, (numVars) := lapply(.SD, as.numeric), .SDcols = numVars]

   data <- data[ data$type == 3, ]
   data[, type := NULL]

   data[, rawTime := time]
   data[, time := time/1000]
   class(data$time) = c("POSIXt", "POSIXct")

   # drop scanMac & posZ
   data[, `:=`(scanMac = NULL, posZ = NULL)]

   data$angle = roundOrientation(data$orientation)

   data$channel = NULL

   data$posXY <- paste(data$posX, data$posY, sep = "-")

   return(data[mac %in% submacs])
}

#### Test Pre-processor

```{r}
lines <- processLine(raw_offline[4:20])
lines

offline_valid <- raw_offline[validLines(raw_offline)]

head(offline_valid)

length(offline_valid)
```

```{r}
tmp <- lapply(offline_valid[1:17], processLine)

sapply(tmp, nrow)
```

#### Dry Run

```{r}
offline_test <- as.data.table(do.call("rbind", tmp))

offline_test
```

#### Process

```{r}
# Enter Debug Context

# options(error = recover, warn = 2)

offline_data <- lapply(offline_valid, processLine)
offline <- as.data.table(do.call("rbind", offline_data),
                         stringsAsFactors = F)

names(offline) <- c("time", "scanMac", "posX", "posY", "posZ",
                    "orientation", "mac", "signal", 
                    "channel", "type")

numVars <- c("time", "posX", "posY", "posZ",
             "orientation", "signal")

dim(offline)

offline[, (numVars) := lapply(.SD, as.numeric), .SDcols = numVars]

offline <- offline[ offline$type == 3, ]
offline[, type := NULL]

offline[, rawTime := time]
offline[, time := time/1000]
class(offline$time) = c("POSIXt", "POSIXct")

# options(error = recover, warn = 1)
```

#### Data Cleaning

```{r}
summary(offline)
```

### Orientation Exploration

```{r}
length(unique(offline$orientation))
```

```{r, fig.height=5}
plot(ecdf(offline$orientation))
```

```{r}
ggplot(offline, aes(orientation)) +
   stat_ecdf() +
   labs("Orientation")
```

```{r}
ggplot(offline, aes(orientation, fill = ..count..)) +
   geom_histogram(bins = 30) +
   scale_y_continuous(labels = comma) +
   labs("Orientation Value Clusters")
```

offline$angle <- roundOrientation(offline$orientation)

# angle = cleaned orientation column
ggplot(offline, aes(angle)) +
   stat_ecdf()

ggplot(offline, aes(angle, fill = ..count..)) +
   geom_histogram(bins = 30) +
   scale_y_continuous(labels = comma) +
   labs("Cleaned Up Angles")
```

```{r}
with(offline, boxplot(orientation ~ angle,
                      xlab = "nearest 45 degree angle",
                      ylab = "orientation"))

ggplot(offline, aes(angle, orientation, group = angle)) +
   geom_boxplot() +
   labs(title = "Cleaned Angle vs Raw Orientation", 
        x = "nearest 45 degree angle", 
        y = "orientation")
```

### MAC Address

```{r}
c(length(unique(offline$mac)), length(unique(offline$channel)))

table(offline$mac)

ggplot(offline, aes(mac, fill = ..count..)) +
   geom_histogram(stat = "count", bins = 30) +
   scale_y_continuous(labels = comma)

# keep only the top 7 MAC address data points

dim(offline)

offline_macs <- names(sort(table(offline$mac), decreasing = T)[1:7])

dim(offline)
```

```{r}
macChannel <- with(offline, table(mac, channel))
apply(macChannel, 1, function(x) sum(x > 0))

# mac and channel are 1:1, we can remove channel
#offline$channel := NULL
```

### Exploring the Position of the Hand-Held Device

```{r}
locDF <- with(offline,
              by(offline, list(posX, posY), function(x) x))

length(locDF)

sum(sapply(locDF, is.null))

locDF <- locDF[ !sapply(locDF, is.null)]

length(locDF)

locCounts <- sapply(locDF, nrow)

locCounts <- sapply(locDF,
                    function(df)
                       c(df[1, c("posX", "posY")], count = nrow(df)))

class(locCounts)

dim(locCounts)

locCounts[ , 1:8]

locCountsDF <- as.data.table(t(locCounts))

locCountsDF$posX <- unlist(locCountsDF$posX)
locCountsDF$posY <- unlist(locCountsDF$posY)

ggplot(locCountsDF, aes(posX, posY, label = count)) +
   geom_text(angle = 45) +
   labs(title = "X / Y postion counts")
```

### Final Data Prep

```{r}
offline <- readData(file_offline, offline_macs)
```

#### Signal Strength

```{r, fig.height=7}
bwplot(signal ~ factor(angle) | mac, data = offline,
       subset = posX == 2 & posY == 12 & 
          mac != "00:0f:a3:39:dd:cd",
       layout = c(2, 3))

summary(offline$signal)
```

```{r, fig.height=7}
densityplot(~ signal | mac + factor(angle), data = offline,
            subset = posX == 24 & posY == 4 &
               mac != "00:0f:a3:39:dd:cd",
            bw = 0.5, plot.points = F)
```

```{r, fig.width=5}
byLocAngleAP <- with(offline,
                     by(offline, list(posXY, angle, mac),
                        function(x) x))
signalSummary <-
   lapply(byLocAngleAP,
          function(oneLoc) {
             ans = oneLoc[1, ]
             ans$medSignal = median(oneLoc$signal)
             ans$avgSignal = mean(oneLoc$signal)
             ans$num = length(oneLoc$signal)
             ans$sdSignal = sd(oneLoc$signal)
             ans$iqrSignal = IQR(oneLoc$signal)
             ans
          })

offlineSummary <- do.call("rbind", signalSummary)

breaks <- seq(-90, -30, by = 5)

bwplot(sdSignal ~ cut(avgSignal, breaks = breaks),
       data = offlineSummary,
       subset = mac != "00:0f:a3:39:dd:cd",
       xlab = "Mean Signal", ylab = "SD Signal")
```

```{r}
with(offlineSummary,
     smoothScatter((avgSignal - medSignal) ~ num,
                   xlab = "Number of Observations",
                   ylab = "mean - median"))
abline(h = 0, col = "#984ea3", lwd = 2)

lo.obj <- with(offlineSummary,
               loess(diff ~ num,
                     data = data.frame(diff = (avgSignal - medSignal),
                                       num = num)))

lo.obj.pr <- predict(lo.obj, newdata = data.frame(num = (70:120)))
lines(x = 70:120, y = lo.obj.pr, col = "#4daf4a", lwd = 2)
```

#### Signal and Distance

```{r, fig.height=6}

subMacs <- names(sort(table(offline$mac), decreasing = T))[1:7]

surfaceSS <- function(data, mac, angle) {
   oneAPAngle <- subset(offlineSummary, mac == mac & angle == angle)

   smothSS <- Tps(oneAPAngle[, c("posX", "posY")],
                  oneAPAngle$avgSignal)

   vizSmooth <- predictSurface(smothSS)

   plot.surface(vizSmooth, type = "C")
   points(oneAPAngle$posX, oneAPAngle$posY, pch=19, cex = 0.5)
}

surfaceSS(offlineSummary, subMacs[5], 0)
```

```{r, fig.height = 7}
parCur <- par(mfrow = c(2,2), mar = rep(1, 4))
mapply(surfaceSS, mac = subMacs[ rep(c(5, 1), each = 2)],
       data = list(data = offlineSummary))
par(parCur)
```

Exclude one of two similar access points

```{r}
offlineSummary <- subset(offlineSummary, mac != subMacs[2])
```

```{r}
AP <- matrix( c( 7.5, 6.3, 2.5, -.8, 12.8, -2.8,
                 1, 14, 33.5, 9.3, 33.5, 2.8),
              ncol = 2, byrow = T,
              dimnames = list(subMacs[ -2 ], c("x", "y") ))

ggplot(data.table(mac = rownames(AP), AP), aes(x, y)) +
   geom_point()
```

```{r, fig.height=7}
diffs <- offlineSummary[, c("posX", "posY")] - AP[offlineSummary$mac, ]

offlineSummary$dist <- sqrt(diffs[, 1]^2 + diffs[, 2]^2)

xyplot(signal ~ dist | factor(mac) + factor(angle),
       data = offlineSummary, pch = 19, cex = 0.3,
       xlab = "distance")
```

### Nearest Neighbor Methods to Predict Location

```{r}
macs <- unique(offlineSummary$mac)

online <- readData(file_online, submacs = macs)

length(unique(online$posXY))
```

```{r}
tabonlineXYA = table(online$posXY, online$angle)
tabonlineXYA[1:6, ]
```

```{r}
keepVars <- c("posXY", "posX", "posY", "orientation", "angle")

byLoc <- with(online,
              by(online, list(posXY),
                 function(x) {
                    ans <- x[1, ..keepVars]
                    avgSS <- tapply(x$signal, x$mac, mean)
                    y = matrix(avgSS, nrow = 1, ncol = 6,
                               dimnames = list(ans$posXY, names(avgSS)))
                    cbind(ans, y)
                 }))

onlineSummary <- do.call("rbind", byLoc)

dim(onlineSummary)

names(onlineSummary)
```

#### Choice of Orientation

reshapeSS <- function(data, varSignal = "signal",
                      keepVars = c("posXY", "posX", "posY")) {

   byLocation <- with(data,
                 by(data, list(posXY),
                    function(x) {
                       ans <- x[1, ..keepVars]
                       avgSS <- tapply(x$signal, x$mac, mean)
                       y = matrix(avgSS, nrow = 1, ncol = 6,
                                  dimnames = list(ans$posXY, names(avgSS)))
                       cbind(ans, y)
                    }))

   newDataSS <- do.call("rbind", byLocation)

   col_names <- colnames(newDataSS)
   n_cols <- length(col_names)

   colnames(newDataSS)[4:n_cols] <- sapply(col_names[4:n_cols], function(col) {
      n <- nchar(col)
      substr(col, n - 2, n)
   })

   newDataSS[, 4:ncol(newDataSS)] <- round(newDataSS[, 4:ncol(newDataSS)])

   return(newDataSS)
}

selectTrain <- function(angleNewObs, signals, m) {

   refs <- seq(0, by = 45, length = 8)

   nearestAngle <- roundOrientation(angleNewObs)

   if( m %% 2 == 1) {
      angles <- seq(-45 * (m - 1) / 2, 45 * (m - 1) / 2, length = m)
   } else {
      m = m + 1
      angles <- seq(-45 * (m - 1) / 2, 45 * (m - 1) / 2, length = m)

      if( sign(angleNewObs - nearestAngle) >= 1)
         angles = angles[ -1 ]
      else
         angles = angles[ -m ]
   }

   angles <- angles + nearestAngle
   angles[angles < 0] = angles[ angles < 0] + 360
   angles[angles > 360] = angles[ angles > 360 ] - 360

   signals <- signals[angle %in% angles, ]

   reshapeSS(signals, varSignal = "avgSignal")
}

train130 <- selectTrain(130, offlineSummary, m = 3)

dim(train130)

findNN <- function(newSignal, trainSubset) {
   diffs <- apply(trainSubset[ , 4:9], 1,
                  function(x) x - newSignal)
   dists <- apply(diffs, 2, function(x) sqrt(sum(x^2)) )
   closest <- order(dists)
   return(trainSubset[closest, 1:3 ])
}

predXY <- function(newSignals, newAngles, trainData,
                   numAngles = 1, k = 3) {

   closeXY <- list(length = nrow(newSignals))

   for(i in 1:nrow(newSignals)) {
      trainSS <- selectTrain(newAngles[i], trainData, m = numAngles)
      closeXY[[i]] = 
         findNN(newSignal = as.numeric(newSignals[i, ]), trainSS)
   }

   estXY = lapply(closeXY,
                  function(x) sapply(x[, 2:3],
                                     function(x) mean(x[1:k])))

   estXY <- do.call("rbind", estXY)

   return(estXY)
}

estXYk1 <- predXY(newSignals = onlineSummary[, 6:11],
                  newAngles = onlineSummary[, 4],
                  offlineSummary, numAngles = 3, k = 1)

estXYk3 <- predXY(newSignals = onlineSummary[, 6:11],
                  newAngles = onlineSummary[, 4],
                  offlineSummary, numAngles = 3, k = 3)
```

```{r}
calcError <- function(estXY, actualXY)
   sum( rowSums( (estXY - actualXY) ^ 2) )

actualXY <- onlineSummary[, c("posX", "posY")]
sapply(list(estXYk1, estXYk3), calcError, actualXY)

>>>>

---
title: ''
mainfont: Arial
fontsize: 12pt
documentclass: report
header-includes:
- \PassOptionsToPackage{table}{xcolor}
- \usepackage{caption}
- \usepackage{amssymb}
- \usepackage{booktabs}
- \usepackage{longtable}
- \usepackage{array}
- \usepackage{multirow}
- \usepackage{wrapfig}
- \usepackage{float}
- \usepackage{colortbl}
- \usepackage{pdflscape}
- \usepackage{tabu}
- \usepackage{threeparttable}
- \usepackage{threeparttablex}
- \usepackage[normalem]{ulem}
- \usepackage{makecell}
- \usepackage[table]{xcolor}
- \usepackage{fancyhdr}
- \usepackage{boldline}
- \usepackage{tipa}
   \definecolor{headergrey}{HTML}{545454}
   \definecolor{msdblue}{HTML}{1C93D1}
   \pagestyle{fancy}
   \setlength\headheight{30pt}
   \rhead{\color{headergrey}\today}
   \fancyhead[L]{\color{headergrey}Moretz, Brandon}
   \fancyhead[C]{\Large\bfseries\color{headergrey}Intoduction to Numerical Analysis}
   \rfoot{\color{headergrey}Chapter 1}
   \lfoot{\color{headergrey}}
   \fancyfoot[C]{\rmfamily\color{headergrey}Computational Methods for Numerical Analysis}
geometry: left = 1cm, right = 1cm, top = 2cm, bottom = 3cm
date: "`r format(Sys.time(), '%d %B, %Y')`"
output:
  pdf_document:
    fig_caption: yes
    latex_engine: xelatex
editor_options: 
  chunk_output_type: console
---

@@

library(cmna, quietly = TRUE, warn.conflicts = FALSE)

@@

## Introduction to Numerical Analysis

### Numerical Analysis in R

Example, factorial:

```{r}
factorial <- function(n) {
   if(n == 0)
      return(0)
   if(n == 1)
      return(1)
   return(factorial(n - 1) + factorial(n - 2))
}

sapply(1:10, factorial)
```

```{r}
isPrime <- function(n) {
   if( n == 2)
      return(TRUE)

   for(i in 2:sqrt(n))
      if(n %% i == 0)
         return(FALSE)

   return(TRUE)
}

nums <- 1:25

rbind(nums, sapply(nums, isPrime))
```

```{r}
x <- c(TRUE, FALSE, TRUE, FALSE, TRUE)
sum(x); mean(x); length(x)
```

```{r}
naievesum <- function(x) {

   s <- 0
   n <- length(x)

   for(i in 1:n) {
      s <- S + x[i]
   }

   s
}

naivesum(c(1, 2, 3, 4.5, -6))
```

```{r}
pwisesum <- function(x) {

   n <- length(x)

   if(n == 1)
      return(x)

   m = floor(n/2)

   return(pwisesum(x[1:m]) + pwisesum(x[(m + 1):n]))
}

pwisesum(c(1, 2, 3, 4.5, -6))

```{r}
kahan_sum <- function(x) {
   comp <- s <- 0
   n <- length(x)

   for(i in 1:n) {
      y <- x[i] - comp
      t <- x[i] + s
      comp <- (t - s) - y
      s <- t
   }

   return(s)
}

kahan_sum(1:8)
```

```{r}
naive_poly <- function(x, coefs) {
   y <- rep(0, length(x))

   for(i in 1:length(coefs)) {
      y <- y + coefs[i] * (x ^ (i - 1))
   }

   return(y)
}

# example poly:

# f(x) = (x-5)(x+3)(x+2)(x-1)
# = x^4 + x^3 - x^2 - 11x + 30
# = 30 - 11 x - 19 x^2 - x^3 + x^4

f <- c(30, -11, -19, -1, 1)

x <- c(-1, 0, 1)

naive_poly(x, f)

```{r}
better_poly <- function(x, coefs) {
   y <- rep(0, length(x))

   cached.x <- 1

   for(i in 1:length(coefs)) {
      y <- y + coefs[i] * cached.x
      cached.x <- cached.x * x
   }

   return(y)
}

better_poly(x, f)
```

```{r}
horner <- function(x, coefs) {
   y <- rep(0, length(x))

   for(i in length(coefs):1)
      y <- coefs[i] + x * y

   return(y)
}

horner(x, f)
```

```{r}
nth_root <- function(a, n, tol = 1 / 1000) {
   x <- 1
   deltax <- tol * 10

   while(abs(deltax) > tol) {
      deltax <- (1 / n) * (a / x ^ (n - 1) - x)
      x <- x + deltax
   }

   return(x)
}

nth_root(100, 2)
nth_root(65536, 4)
nth_root(1000, 3)
nth_root(pi, 2)
```

## Exercises

#### 1.)

What is the circumference of a 10-meter-diameter circle symbolically?

$2\pi r$, r = 10

$10\pi$

What is it numerically?

```{r}
10*pi
```

Which one is correct? *Both*

#### 2.)

Describe an application for which a numerical solution are preferable to a symbolic solution.

_An exact symbolic solution to some problems can be difficult or near impossible, so a good approximation via numerical methods is preferable than no solution._

Describe an application where symbolic solutions are preferable to numerical.

_An analytical/symbolic solution can solve a related family of problems._

#### 3.)

If an unsigned variable were stored using 16 bits, what is the largest number it could hold?

```{r}
2^16
```

#### 4.)

Using R's internal trigonometric functions, implement a secant function.

```{r}
sec <- function(x) 1/cos(x)

assertthat::are_equal(sec(pi), -1)

#### 5.)

Using the exponentiation operator, create a function that finds the nth root of any number.

```{r}
nth_root <- function(x, n) x^(1/n)

assertthat::are_equal(nth_root(32, 5), 2)

#### 6.)

Function 1.1 provides a recursive implementation of the Fibonacci sequence. Rewrite the function to provide a result without recursion.

```{r}
fib_iter <- function(x) {

   # build look-up
   fib_values <- list()
   fib_values[[1]] <- 1

   for(index in 2:x) {

      fib1 <- ifelse(index <= 1, 1, fib_values[[index - 1]])
      fib2 <- ifelse(index <= 1, 1, ifelse(index == 2, 1, fib_values[[index - 2]]))

      fib_values[index] <- fib1 + fib2
   }

   return(fib_values[[x]])
}

sapply(1:10, fib_iter)
```

What are the relative strenghts and weaknesses of each implementation?

_The recursive version is a bit more readable._

#### 7.)

Implement a product function in R without using recursion.

@@

@@

#### 8.)
#### 9.)
#### 10.)
#### 11.)
#### 12.)
#### 13.)
#### 14.)
#### 15.)
#### 16.)
#### 17.)
#### 18.)
#### 19.)

### 2020-04-06 Mathematics and Statistics for FRM/03_Statistics.rmd
sln refactor + case studies/computational methods/statistical computing work
sln refactor + case studies/computational methods/statistical computing work

---
title: ''
mainfont: Arial
fontsize: 12pt
documentclass: report
header-includes:
- \PassOptionsToPackage{table}{xcolor}
- \usepackage{caption}
- \usepackage{amssymb}
- \usepackage{booktabs}
- \usepackage{longtable}
- \usepackage{array}
- \usepackage{multirow}
- \usepackage{wrapfig}
- \usepackage{float}
- \usepackage{colortbl}
- \usepackage{pdflscape}
- \usepackage{tabu}
- \usepackage{threeparttable}
- \usepackage{threeparttablex}
- \usepackage[normalem]{ulem}
- \usepackage{makecell}
- \usepackage[table]{xcolor}
- \usepackage{fancyhdr}
- \usepackage{boldline}
- \usepackage{tipa}
   \definecolor{headergrey}{HTML}{545454}
   \definecolor{msdblue}{HTML}{1C93D1}
   \pagestyle{fancy}
   \setlength\headheight{30pt}
   \rhead{\color{headergrey}\today}
   \fancyhead[L]{\color{headergrey}Moretz, Brandon}
   \fancyhead[C]{\Large\bfseries\color{headergrey}Basics}
   \rfoot{\color{headergrey}Chapter 1}
   \lfoot{\color{headergrey}}
   \fancyfoot[C]{\rmfamily\color{headergrey}Mathematics and Statistics for Financial Risk Management}
geometry: left = 1cm, right = 1cm, top = 2cm, bottom = 3cm
date: "`r format(Sys.time(), '%d %B, %Y')`"
output:
  pdf_document:
    fig_caption: yes
    latex_engine: xelatex
editor_options: 
  chunk_output_type: console
---

@@

#### Chapter 1

@@

Solve for $y$, where:

a.) $y = ln(e^5)$

5

b.) $y = ln(1/e)$

$y = ln(1/e) = ln(1) - 1 = 0 - 1 = -1$

c.) $y = ln(10e)$

$y = ln(10e) = ln(10) + ln(e) = ln(10) + 1$

```{r}
log(10) + 1
```

@@

The nominal monthly rate for a loan is quoted at 5%. What is the equivalent annual rate? Semiannual rate? Continious rate?

```{r}
effective_rate <- function(rate, periods) {
   ((1 + rate/periods)^periods - 1)
}
```

Annual:

```{r}
effective_rate(.05, 12) * 100
```

Semiannual:

```{r}
effective_rate(.05, 2) * 100
```

Continous:

```{r}
log(1 + effective_rate(.05, 12)) * 100
```

@@

Over the course of a year, the log return on a stock market index is 11.2%. The starting value of the index is 100. What is the value at the end of the year?

```{r}
100 * exp(.112)
```

@@

You have a portfolio of 10 bonds. In how many different ways can exactly two bonds default? Assume the order in which the bonds default is unimportant.

```{r}
choose(10, 2)
```

@@

What is the present value of a perpetuity that pays $100 per year? Use an actual discount rate of 4%, and assume the first payment will be made in exactly one year.

```{r}
100 * (1/0.04)
```

@@

ABC stock will pay a $1 dividend in one year. Assume the dividend will continue to be paid annually forever and the dividend payments will increase in size at a rate of 5%. Value this stream of dividends using a 6% annual discount rate.

```{r}
1 / (0.06 - 0.05)
```

@@

What is the present value of a 10-year bond with a $100 face value, which pays a 6% coupon annually? Use an 8% discount rate.

@@

present_value <- function(face, int.rate, periods, discount.rate) {
   delta <- 1/(1 + discount.rate)

   coupon <- face * int.rate

   discount <- (delta - delta^(periods+1)) / (1 - delta)

   coupon * discount + ( face / (1 + discount.rate)^periods )
}

present_value(100, .06, 10, .08)
```

@@

Solve for $x$, where $e^{e^x} = 10$

```{r}
log(log(10))
```

@@

Calculate the value of the following summation: $\sum^{9}_{i=0}(-0.5)^i$

```{r}
r <- -0.5; n <- 9

(1 - r^n) / (1 - r)
```

@@

The risk department of your firm has 10 analysts. You need to select four analysts to serve on a special audit committee. How many possible groupings of four analysts can be put together?

```{r}
choose(10, 4)
```

long-hand:

```{r}
n <- 10; k <- 4

factorial(10) / ( factorial(k) * factorial(n - k))
```

@@

What is the present value of a newly issued 10-year bond with a notial value of $100 and a 2% annual coupon? Assume a constant 5% annual discount rate and no risk of default.

```{r}
present_value(100, .02, 10, 0.05)
```

>>>>

---
title: ''
mainfont: Arial
fontsize: 12pt
documentclass: report
header-includes:
- \PassOptionsToPackage{table}{xcolor}
- \usepackage{caption}
- \usepackage{amssymb}
- \usepackage{booktabs}
- \usepackage{longtable}
- \usepackage{array}
- \usepackage{multirow}
- \usepackage{wrapfig}
- \usepackage{float}
- \usepackage{colortbl}
- \usepackage{pdflscape}
- \usepackage{tabu}
- \usepackage{threeparttable}
- \usepackage{threeparttablex}
- \usepackage[normalem]{ulem}
- \usepackage{makecell}
- \usepackage[table]{xcolor}
- \usepackage{fancyhdr}
- \usepackage{boldline}
- \usepackage{tipa}
   \definecolor{headergrey}{HTML}{545454}
   \definecolor{msdblue}{HTML}{1C93D1}
   \pagestyle{fancy}
   \setlength\headheight{30pt}
   \rhead{\color{headergrey}\today}
   \fancyhead[L]{\color{headergrey}Moretz, Brandon}
   \fancyhead[C]{\Large\bfseries\color{headergrey}Statistics}
   \rfoot{\color{headergrey}Chapter 3}
   \lfoot{\color{headergrey}}
   \fancyfoot[C]{\rmfamily\color{headergrey}Mathematics and Statistics for Financial Risk Management}
geometry: left = 1cm, right = 1cm, top = 2cm, bottom = 3cm
date: "`r format(Sys.time(), '%d %B, %Y')`"
output:
  pdf_document:
    fig_caption: yes
    latex_engine: xelatex
editor_options: 
  chunk_output_type: console
---

@@

#### Chapter 3

@@

Compute the mean and the median of the following series of returns:

```{r}
dat <- data.table(Returns = c(.12, .5, -.08, .2, .04, .1, .02))

pretty_kable(dat, "Returns")
```

```{r}
mean(dat$Returns)

median(dat$Returns)
```

@@

Compute the sample mean and the standard deviation of the following returns:

```{r}
dat <- data.table(Returns = c(.12, .5, -.08, .2, .04, .1, .02))

pretty_kable(dat, "Returns")
```

@@

Prove that Equation 3.2 is an unbiased estimator of the mean. That is, show that $\mathbb{E}[\hat\mu] = \mu$

$\mathbb{E}(\bar{x}) = \mathbb{E}[\frac{x_1 + x_2 + x_3 + \ldots + x_n}{n}]$

$\ldots = \mathbb{E}\frac{1}{n}(x_1 + x_2 + x_3 + \ldots + x_n)$

$\ldots = \frac{1}{n}\mathbb{E}(x_1 + x_2 + x_3 + \ldots + x_n)$

$\ldots = \frac{1}{n}(\mathbb{E}[x_1] + \mathbb{E}[x_2] + \mathbb{E}[x_3] + \ldots + \mathbb{E}[x_n])$

$\ldots = \frac{1}{n}(\mu_1 + \mu_2 + \mu_3 + \ldots + \mu_n)$

$\ldots = \frac{1}{n}n(u)$

$\ldots = \mu$

@@

What is the standard deviation of the estimator in Eq. 3.2? Assume the various data points are i.i.d.

$\mu = \frac{1}{n}\sum^{n}_{i=1}r_i$

$\frac{1}{n}\sum^{n}_{i=1}r_i = \frac{1}{n}(r_1 + r_2 + r_3 \ldots + r_n)$

$\sigma^2 = \frac{1}{n}\sum(x_i - \mu)^2$

$\sigma_{\mu} = \frac{\sigma_{}}{\sqrt{n}}$

@@

Calculate the population covariance and correlation of the following series:

```{r}
dat <- data.table(S1 = c(.21, .53, .83, .19), S2 = c(.2, .32, .8, .4))

pretty_kable(dat, "Returns")
```

```{r}
cov(dat$S1, dat$S2)

cor(dat$S1, dat$S2)
```

@@

Calculate the population mean, standard deviation and skewness of each of the following two series:

```{r}
dat <- data.table(S1 = c(-51, -21, 21, 51), S2 = c(-61, -7, 33, 35))

pretty_kable(dat, "Returns")
```

skew <- function(x) {
   mean(x - mean(x))^3 / sd(x)^3
}

apply(dat, 2, mean)
apply(dat, 2, sd)
apply(dat, 2, skew)

@@

Calculate the population mean, standard deviation and skewness of each of the following two series:

```{r}
dat <- data.table(S1 = c(-23, -7, 7, 23), S2 = c(-17, -17, 17, 17))

pretty_kable(dat, "Returns")
```

@@

skew <- function(x) {
   mean(x - mean(x))^3 / sd(x)^3   
}

apply(dat, 2, mean)
apply(dat, 2, sd)
apply(dat, 2, skew)

@@

Given the probability density function for a random variable, X

$f(x) = \frac{x}{18}$ for $0 \le x \le 6$

$\mu = \int_0^6 \frac{x^3}{18 * 3} = \frac{6^3}{3*18} - \frac{0^3}{3*18}$

$\frac{6^2}{3^2} = 4$

$\sigma^2 = \int_0^6(x - 4)^2\frac{x}{18}dx$

$\frac{1}{18}(x^3 - 8x^2 + 16x)dx$

$\sigma^2 = 2(9 - 16 + 8) = 2$

@@

Prove that Equation 3.19, reproduced here, is an unbiased estimator of the variance.

$\hat\sigma^2_x = \frac{1}{n - 1}\sum_{i=1}^n(x_i - \hat\mu_x)$

$\hat\sigma^2_x = \frac{1}{n - 1}\sum_{i=1}^n = \frac{1}{n - 1}\sum_{i=1}^n(\frac{n-1}{n}x_i - \frac{1}{n}\sum_{j \ne x}x_j)^2$

$\hat\sigma^2_x = \frac{1}{n}\sum_{i=1}^nx^2_i - \frac{1}{n(n - 1)}\sum{i=1}^n\sum_{i \ne j}x_ix_j$

$\mathbb{E}[x^2_i] = \sigma^2 + \mu^2$

$\mathbb{E}[x_i x_j] = \mu_i\mu_j + iff \ \sigma_{ij} = 0 \ \forall i \ne j$

@@

Given two random variables, $X_A$ and $X_B$, with the corresponding means $\mu_A$ and $\mu_B$ and standard deviations $\sigma_A$ and $\sigma_B$, prove that the variance of $X_A$ plus $X_B$ is:

$\mathbb{Var}[X_A + X_B] = \sigma^2_A + \sigma^2_B + 2\rho_{AB}\sigma_A\sigma_B$

$\mathbb{E}[X_A + X_B] = \mathbb{E}[X_A] + \mathbb{E}[X_B] = \mu_A + \mu_B$

$\mathbb{Var}[X_A + X_B] = \mathbb{E}[(X_A + X_B - \mathbb{E}[X_A + X_b])^2]$

$\mathbb{Var}[X_A + X_B] = \sigma^2_A + \sigma^2_B + 2Cov[X_A, X_B]$

@@

A $100 notional, zero coupon bond has one year to expiry. The probability of default is 10%. In the event of default, assume that the recovery rate is 40%.

The continiously compounded discount rate is 5%. What is the present value of this bond?

```{r}
EV <- .9 * 100 + 0.10 * 40
```

$\mathbb{E}(V) = `r EV`$

```{r}
PV <- exp(-.05) * EV
```

$\mathbb{P}(V)$ = $`r round(PV, 2)`
